---
title: "Inspect available gauge data"
format: html
editor: visual
---

```{r}
# Using my fork of buzacott/bomWater
# devtools::install_github('callum-savage/bomWater')
library(bomWater)
library(tidyverse)
```

## Overview

This is an exploratory notebook to see what data is available from gauges along the Narran River.

-   What timeseries variables are available from river gauges along the Narran River?

-   When does each time series start and end?

-   How much data is missing?

-   Is the data reliable?

-   Are the pre-aggregated time-series reliable?

## Get station details

### Identify relevant stations

I am limiting the scope initially to the Narran River, and gauges directly upstream. From Bom Water Data Online, we can see the following stations.

-   The Balonne river flows south through St George, QLD
-   20 km North of Dirranbandi, the Culgoa River branches off to the west of the Balonne (and away from the Narran Lakes)
    -   There are two gauges here, and it looks like both may be relevant
        -   422204A, Culgoa River at Whyenbah
        -   422205B, Balonne Minor River at Hastings
-   South of Dirranbandi, the Balonne River branches into the Narran River (eastern) and Bokahra River (western)
    -   This branching is ungaguged, so the northernmost gauge on the Bokhara River might be important
        -   422209A, Bokhara River at Hebel
-   The Narran River passes into NSW terminates at the Narran Wetlands
    -   QLD
        -   422206A, Narran River at Dirranbandi-Hebel Road
    -   NSW
        -   422016, NARRAN RIVER AT WILBY WILBY (BELVEDERE)
        -   422030, NARRAN RIVER AT NEW ANGLEDOOL NO. 2
    -   Narran lakes
        -   422013, NARRAN RIVER \@ BUNDAH
        -   422034, NARRAN RIVER \@ BACK LAKE,
        -   422029, NARRAN RIVER \@ NARRAN PARK

### Check station measurements

See which variables were measured at each station

```{r}
station_numbers <- c("422204A", "422205B", "422209A", "422206A", "422016",
                     "422030", "422013", "422034", "422029")

station_parameters <- get_parameter_list(
    station_number = station_numbers,
    return_fields = c(
      "station_no",
      "parametertype_name",
      "parametertype_unitname")
  ) |> 
  rename(
    parameter = parametertype_name, 
    parameter_unit = parametertype_unitname
  )

station_parameters
```

Not all stations have each parameter, so create a table explicitly showing which measurements are available at which stations.

```{r}
# Create a function to shorten parameter names
shorten_parameter <- function(long_param) {
  long_param |>
    str_remove(" @ 25C") |>
    str_extract("\\w+$") |> # last word
    str_to_lower()
}

station_measurements <- station_parameters |>
  mutate(parameter_short = shorten_parameter(parameter)) |>
  mutate(measured = TRUE) |>
  pivot_wider(
    id_cols = station_no,
    names_from = parameter_short,
    values_from = measured,
    values_fill = FALSE
  )

station_measurements
```

-   All stations measure water level

-   Only one station (Back Lake) is missing discharge data

-   Other variables are a bit more spotty - especially pH which is only measured at one station
